Job mobility and earnings growth: diffrerent markets, different outcomes.

### 2018-09-07 draft.Rmd
Update draft.Rmd

Countries differ in their rate of job mobility and turnover. Researchers cite employment protection legislation as the key driver to such rates; however, the rewards or effects tied to mobility may play a stronger role in determining these differences. We consider this idea through two aims. First we compare the effects of quits and promotions on earnings growth. Second, we consider their effects in two institutional settings (Germany and the UK). Studies exploring the effects of job mobility are rare, studies which compare and contrast their effects in separate institutional frameworks, or capitalisms, are rarer. We offer two findings. First, quits and promotions have significantly different effects on earnings growth, depsite both commonly labeled as "voluntary". Internal promotions yield significant earnings growth for workers in both countries. Quits have a minor effect, which is explained by a change in weekly working hours. Second, the effect of promotions is larger in Germany than in the UK. This suggests the premium tied to promotion is greater in Germany than the UK. The high premium tied to internal mobility in Germany may act as an incentive to commit to the firm instead of the market, hence the lower rate of mobility among German workers. The article helps to understand country-differences in mobility rates; our explanations go beyond a nation's level of employment protection.

@@

Worker mobility is a key strategy in improving outcomes like wages and conditions. Routinely, authors show that workers stand to gain from mobility within and between firms. Despite this, capitalist countries differ in their rates of job mobility. This poses the question, are workers in less mobile countries starved for vacancies or opprtunities? We consider the impact mobility has on internal and external job mobility for the core workforce

>>>>

Countries differ in their rate of job mobility and turnover. Researchers cite employment protection legislation as key to understanding the rates; however, the rewards or effects tied to mobility play a stronger role in determining differences. We consider this idea through two aims. First we compare the effects of quits and promotions on earnings growth. Second, we consider their effects in two institutional settings (Germany and the UK). Studies exploring the effects of job mobility are rare, studies which compare and contrast their effects in separate institutional frameworks, or capitalisms, are rarer. We offer two findings. First, quits and promotions have significantly different effects on earnings growth, depsite both commonly labeled as "voluntary". Internal promotions yield significant earnings growth for workers in both countries. Quits have a minor effect, which is explained by a change in weekly working hours. Second, the effect of promotions is larger in Germany than in the UK. This suggests the premium tied to promotion is greater in Germany than the UK. The high premium tied to internal mobility in Germany may act as an incentive to commit to the firm instead of the market, hence the lower rate of mobility among German workers. The article helps to understand country-differences in mobility rates; our explanations go beyond a nation's level of employment protection.

@@

Worker mobility is a key strategy in improving outcomes like wages and conditions. Routinely, authors show that workers make significant gains through quitting jobs or securing promotions. Despite this, capitalist countries differ by job mobility rates, specifically in their rates of voluntary job mobility. Authors typically explain these differences as a difference in the hiring and firing of workers (Brown) or a difference in the rate of job creation and destruction (OECD). This article argues that mobility rates also differ due to the rewards they offer to workers. More specifically, we propose that countries differ in the rewards tied to quits and promotions.

Institutional differences in rewarding mobility rarely appear in theoretical literature. In the *job-matching* model mobility stems from a mismatch between a worker's skills, and her rewards or conditions (Jovanvich). In this way countries with low rates of job mobility, are simply countries with a good record of job matching. In the *job-search* model mobility stems from the opportunity of higher wages. Thus if labour markets have a wide dispersion of wages workers have more opportunity to match their reservation wages, which leads to more mobility. *Human capital* theorists would see mobile labour markets as those with a high dependence on transferable skills, and less mobile countries as those dependent on firm- specific skills.

Traditionally, authors have focused extensively on what drives workers to quit. This approach often took for granted that job mobility is of interest to workers, rewarding them under certain conditions.
